### 2024-09-07 base-rates.qmd
20240907 - updates
20240907 - copyedit

$$
\begin{aligned}
  P(\text{Blue}) &= .15 && \text{prior probability of a Blue cab}\\
  P(\text{Green}) &= .85 && \text{prior probability of a Green cab}\\
  P(\text{Correct}|\text{Blue}) &= .80 && \text{probability the witness correctly identifies a Blue cab}\\
  P(\text{Correct}|\text{Green}) &= .80 && \text{probability the witness correctly identifies a Green cab}\\
  P(\text{Incorrect}|\text{Blue}) &= .20 && \text{probability the witness incorrectly identifies a Blue cab}\\
  P(\text{Incorrect}|\text{Green}) &= .20 && \text{probability the witness incorrectly identifies a Green cab}\\
\end{aligned}
$$

@@

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)}\\
  P(\text{Blue}|\text{Identified as Blue}) &= \frac{P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue})}{P(\text{Identified as Blue})}
\end{aligned}
$$

We can compute the term in the denominator ($P(\text{Identified as Blue})$) using the law of total probability (described in @sec-bayesTheorem).
$$
\begin{aligned}
  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i) \\
  P(R_i) &= P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue}) + P(\text{Identified as Blue}|\text{Green}) \cdot P(\text{Green})
  0.29 &= (.80 \times .15) + (.20 \times .85) \\
\end{aligned}
$$

@@

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)}\\
  P(\text{Blue}|\text{Identified as Blue}) &= \frac{P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue})}{P(\text{Identified as Blue})}\\
  0.414 &= \frac{0.80 \times 0.15}{0.29}
\end{aligned}
$$

@@

Thus, there was a 41.4% probability that the car involved in the accident was Blue rather than Green.
However, when faced with this problem, people tend to ignore the base rate and go with the witness [@Kahneman2011].
According to @Kahneman2011, the most frequent response to this question regarding is that there is an 80% that the car was Blue.

>>>>

$$
\begin{aligned}
  P(\text{Blue}) &= .15 && \text{prior probability of a Blue cab}\\
  P(\text{Green}) &= .85 && \text{prior probability of a Green cab}\\
  P(\text{Correct}|\text{Blue}) &= .80 && \text{probability the witness correctly identifies a Blue cab}\\
  P(\text{Correct}|\text{Green}) &= .80 && \text{probability the witness correctly identifies a Green cab}\\
  P(\text{Incorrect}|\text{Blue}) &= .20 && \text{probability the witness incorrectly identifies a Blue cab}\\
  P(\text{Incorrect}|\text{Green}) &= .20 && \text{probability the witness incorrectly identifies a Green cab}\\
\end{aligned}
$$ {#eq-cabExampleGivenInfo}

@@

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)}\\
  P(\text{Blue}|\text{Identified as Blue}) &= \frac{P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue})}{P(\text{Identified as Blue})}
\end{aligned}
$$ {#eq-cabExampleBayesTheorem}

We can compute the term in the denominator ($P(\text{Identified as Blue})$) using the law of total probability (described in @sec-bayesTheorem).
$$
\begin{aligned}
  P(R_i) &= P(R | C) \cdot P(C_i) + P(R | \text{not } C) \cdot P(\text{not } C_i)\\
  P(R_i) &= P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue}) + P(\text{Identified as Blue}|\text{Green}) \cdot P(\text{Green})\\
  0.29 &= (.80 \times .15) + (.20 \times .85) \\
\end{aligned}
$$ {#eq-cabExampleLawOfTotalProbability}

@@

$$
\begin{aligned}
  P(C | R) &= \frac{P(R | C) \cdot P(C_i)}{P(R_i)}\\
  P(\text{Blue}|\text{Identified as Blue}) &= \frac{P(\text{Identified as Blue}|\text{Blue}) \cdot P(\text{Blue})}{P(\text{Identified as Blue})}\\
  0.414 &= \frac{0.80 \times 0.15}{0.29}
\end{aligned}
$$ {#eq-cabExamplePosterior}

@@

Thus, there was a 41.4% probability that the car involved in the accident was Blue rather than Green.
However, when faced with this problem, people tend to [ignore the base rate](#sec-fallaciesBaseRate) and go with the witness [@Kahneman2011].
According to @Kahneman2011, the most frequent response to this question regarding is that there is an 80% that the car was Blue.
